Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}

### 2021-05-13 Latex/main.tex

%% Group authors per affiliation:
\author{Fabian Arzberger, Anton Bredenbeck, Jasper Zevering, Dorit Borrmann, Andreas Nüchter}
\address{Informatics VII: Robotics and Telematics,
Julius-Maximilians-University Würzburg, Germany\\
%One contact suffices
e-mail: dorit.borrmann@uni-wuerzburg.de}

@@

\begin{abstract}
Spherical robots are a format that has not been thoroughly explored for the application of mobile mapping. 
In contrast to other designes it provides some unique advantages. 
Among those are a protective spherical shell which protects internal sensors and actuators from possible harsh environments as well as an inherent rotation for locomotion that enables measurements in all directions.
Mobile mapping always requires a high-precise pose knowledge to obtain consistent and correct environment maps. 
This is typically done by a combination of external reference sensors such as Global Navigation Satellite System (GNSS) measurements and inertial measurements, or by post processing the data and registering the different measurements to each other. 
In indoor environments the GNSS reference is not an option, hence many mobile mapping applications turn to the second option.
An advantage of indoor environments is, that human made environments usually have a certain structure, such as parallel and perpendicular planes.
We propose a registration procedure that exploits this structure by minimizing the distance of measured points to a corresponding plane. 
Further we evaluate the procedure on a simulated dataset of an ideal corridor as well as in experimentally acquired datasets with different motion profiles. 
We show that we nearly reproduce the ground truth for the simulated dataset and improve the average point-to-point distance to a reference scan in the experimental dataset.
The presented algorithms are required to work completely autonomous.
\end{abstract}

>>>>

%% Group authors per affiliation:
\author{Fabian Arzberger and Anton Bredenbeck, Jasper Zevering, Dorit Borrmann, Andreas Nüchter}
\address{Informatics VII: Robotics and Telematics,
Julius-Maximilians-University Würzburg, Germany\\
%One contact suffices
e-mail: dorit.borrmann@uni-wuerzburg.de}

@@

\begin{abstract}
Spherical robots are a format that has not been thoroughly explored for the application of mobile mapping. 
In contrast to other designes, it provides some unique advantages. 
Among those is a protective spherical shell that protects internal sensors and actuators from possible harsh environments, as well as an inherent rotation for locomotion that enables measurements in all directions.
Mobile mapping always requires a high-precise pose knowledge to obtain consistent and correct environment maps. 
This is typically done by a combination of external reference sensors such as Global Navigation Satellite System (GNSS) measurements and inertial measurements or by post processing the data and registering the different measurements to each other. 
In indoor environments, the GNSS reference is not an option.
Hence many mobile mapping applications turn to the second option.
An advantage of indoor environments is that human-made environments usually have a certain structure, such as parallel and perpendicular planes.
We propose a registration procedure that exploits this structure by minimizing the distance of measured points to a corresponding plane. 
Further, we evaluate the procedure on a simulated dataset of an ideal corridor and on an experimentally acquired dataset with different motion profiles. 
We show that we nearly reproduce the ground truth for the simulated dataset and improve the average point-to-point distance to a reference scan in the experimental dataset.
The presented algorithms are required to work completely autonomous.
\end{abstract}

### 2021-05-13 Latex/sections/stateOfTheArt.tex

Up to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before, while there exists camera based approaches~\cite{9233654}.
However, laser-based SLAM algorithms for motions in six degrees of freedom (DoF) have been thoroughly studied.
For outdoor environments~\cite{JFR2006} provides a first baseline. Adding a heuristic for closed loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
Zhang et al. propose a real-time version of the SLAM algorithm in~\cite{Zhang2014}.
They achieve the performance at a lower computational load by dividing the SLAM algorithm into two different algorithms: one performing odometry at a high frequency but low fidelity and another running at a lower frequency performing fine matching and registration of the point clouds.
The more recently Dröschel et al. also proposes an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.

Since these approaches are based on point-to-point correspondences they require rather high point density to achieve precise registration.
For low-cost LiDARs this implies a slow motion and long integration time.
Further, none of the approaches exploit the structure inherent to man-made environments, hence missing out on possible advantages.

@@

The de-facto standard for many SLAM algorithms is the Iterative-Closest-Point (ICP) algorithm~\cite{Besl1992} that employs point-to-point correspondences using closest points, as the name suggests. 
To overcome the requirements on point-density imposed by the point-to-point correspondences instead other correspondences are used. 
In man made environments planes are abundantly available, and hence provide an attractive base for correspondences.
Förster et al. use this property of man made environment successfully in~\cite{Foerstner2017}.
They register point clouds using plane-to-plane correspondences and include uncertainty measures for the detected planes as well as the estimated motion.
Thereby, they propose a costly exact algorithm and cheaper approximations that yield high quality maps.
Favre et al.~\cite{favre2021} use point-to-plane correspondences after preprocessing the point clouds using plane-to-plane correspondences to successfully register two scans with each other.

Both approaches use plane-to-plane correspondences to pre-register the scans. However for pre-registration the classical point-to-point registration is also very effective. 
One advantage that point-to-point correspondences have over plane-to-plane correspondences is, that they do not require a long stop in each pose. For plane-to-plane correspondences this is required to gather enough data to measure planes in each individual scan robustly.
The resulting scan procedure is stop-scan-and-go.
In particular for the application of a spherical robot this standstill in each pose cannot be guaranteed or even approximated making continuous-time approaches using point-to-plane correspondences the method of choice.

In the following we propose this combination of point-to-point based pre-registration followed by a point-to-plane based optimization.

>>>>

Up to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before, while there exists camera-based approaches~\cite{9233654}.
However, laser-based SLAM algorithms for motions in six degrees of freedom (DoF) have been thoroughly studied.
For outdoor environments~\cite{JFR2006} provides a first baseline.
Adding a heuristic for closed-loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
Zhang et al. propose a real-time version of the SLAM algorithm in~\cite{Zhang2014}.
They achieve the performance at a lower computational load by dividing the SLAM algorithm into two different algorithms: one performing odometry at a high frequency but low fidelity and another running at a lower frequency performing fine matching and registration of the point clouds.
More recently Dröschel et al. also propose an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.

Since these approaches are based on point-to-point correspondences, they require a rather high point density to achieve precise registration.
For low-cost LiDARs, this implies slow motion and long integration time.
Further, none of the approaches exploit the structure inherent to human-made environments, hence missing out on possible advantages.

@@

The de-facto standard for many SLAM algorithms is the Iterative-Closest-Point (ICP) algorithm~\cite{Besl1992} that employs point-to-point correspondences using closest points, as the name suggests. 
To overcome the requirements on point-density imposed by the point-to-point correspondences instead other correspondences are used. 
In human-made environments, planes are abundantly available and hence provide an attractive base for correspondences.
Förster et al. use this property of human-made environment successfully in~\cite{Foerstner2017}.
They register point clouds using plane-to-plane correspondences and include uncertainty measures for the detected planes and the estimated motion.
Thereby, they propose a costly exact algorithm and cheaper approximations that yield high-quality maps.
Favre et al.~\cite{favre2021} use point-to-plane correspondences after preprocessing the point clouds using plane-to-plane correspondences to register two scans with each other successfully.

Both approaches use plane-to-plane correspondences to pre-register the scans.
However, for pre-registration the classical point-to-point registration is also very effective. 
One advantage that point-to-point correspondences have over plane-to-plane correspondences is that they do not require a long stop in each pose.
For plane-to-plane correspondences, this is necessary to gather enough data to measure planes in each scan robustly.
The resulting scan procedure is stop-scan-and-go.
In particular, for the application of a spherical robot this standstill in each pose cannot be guaranteed or even approximated, making continuous-time approaches using point-to-plane correspondences the method of choice.

In the following, we propose this combination of point-to-point based pre-registration followed by a point-to-plane based optimization.
